Position: Stop Making Unscientific AGI Performance Claims

### 2024-05-28 abstract.tex
trimmed down abstract slightly

Developments in the field of AI in general, and Large Language Models (LLMs) in particular, have created a `perfect storm’ for observing `sparks’ of Artificial General Intelligence (AGI) that are spurious. Like simpler models, LLMs distill representations in their latent embeddings that have been shown to correlate with meaningful phenomena. Nonetheless, the correlation of such representations has often been linked to human-like intelligence in the latter but not the former. We probe models of varying degrees of sophistication including random projections, matrix decompositions, deep autoencoders and transformers: all of them successfully distill knowledge and yet none of them develop true understanding. Specifically, we show that embeddings of a language model fine-tuned on central bank communications can make meaningful predictions, via correlations with unseen economic variables, such as price inflation. However, we then show that inflation is also predicted for nonsense prompts about growing and shrinking bird populations (`dovelation’). We therefore argue that patterns in latent spaces are spurious sparks of (AGI). Additionally, we review literature from the social sciences that shows that humans are prone to seek patterns and anthropomorphize. We, therefore, argue that both the methodological setup and common public image of AI are ideal for the misinterpretation that correlations between model representations and some variables of interest are 'caused' by the model's understanding of underlying `ground truth’ relationships. We, therefore, call for the academic community to exercise extra caution, and to be keenly aware of principles of academic integrity, in interpreting and communicating about AI research outcomes.

>>>>

Developments in the field of AI in general, and Large Language Models (LLMs) in particular, have created a `perfect storm’ for observing `sparks’ of Artificial General Intelligence (AGI) that are spurious. Like simpler models, LLMs distill meaningful representations in their latent embeddings that have been shown to correlate with external variables. Nonetheless, the correlation of such representations has often been linked to human-like intelligence in the latter but not the former. We probe models of varying complexity including random projections, matrix decompositions, deep autoencoders and transformers: all of them successfully distill information that can be used to predict latent or external variables and yet none of them have ever been linked to AGI. We, therefore, argue that patterns in latent spaces are spurious sparks of AGI. Additionally, we review literature from the social sciences that shows that humans are prone to seek patterns and anthropomorphize. We conclude that both the methodological setup and common public image of AI are ideal for the misinterpretation that correlations between model representations and some variables of interest are `caused' by the model's understanding of underlying `ground truth’ relationships. We, therefore, call for the academic community to exercise extra caution, and to be keenly aware of principles of academic integrity, in interpreting and communicating about AI research outcomes.

### 2024-05-28 body.tex
trimmed down abstract slightly

Noticing these trends, we as the authors of this article are concerned. We feel that the current culture of racing toward Big Outcome Statements in industry and academic publishing too much disincentivizes efforts toward more thorough and nuanced actual problem understanding. At the same time, as the outside world is so eager to adopt AI technology, (too) strong claims make for good sales pitches, but a question is whether there is indeed sufficient evidence for these claims. With successful A(G)I outcomes needing to look human-like, this also directly plays into risks of anthropomorphizing (the attribution of human-like qualities to non-human objects) and confirmation bias (the seeking-out and/or biased interpretation of evidence in support of one's beliefs). In other words, it is very tempting to claim surprising human-like achievements of AI, and as humans, we are very prone to genuinely believing this. \textbf{We therefore urge our fellow researchers to stop making unscientific AGI performance claims}.

>>>>

Noticing these trends, we as the authors of this article are concerned. We feel that the current culture of racing toward Big Outcome Statements in industry and academic publishing too much disincentivizes efforts toward more thorough and nuanced actual problem understanding. At the same time, as the outside world is so eager to adopt AI technology, (too) strong claims make for good sales pitches, but a question is whether there is indeed sufficient evidence for these claims. With successful AGI outcomes needing to look human-like, this also directly plays into risks of anthropomorphizing (the attribution of human-like qualities to non-human objects) and confirmation bias (the seeking-out and/or biased interpretation of evidence in support of one's beliefs). In other words, it is very tempting to claim surprising human-like achievements of AI, and as humans, we are very prone to genuinely believing this. \textbf{We therefore urge our fellow researchers to stop making unscientific AGI performance claims}.
